---
title: "Data Visualization Mini-Project 1"
author: "Dhrishya Menon - dmenon2337@floridapoly.edu "
date: '2022-06-05'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The Dataset chosen for the Mini Project is "1984-2017 fuel economy data produced during vehicle testing at the Environmental Protection Agency’s (EPA) National Vehicle and Fuel Emissions Laboratory"
[Fuel Data ](https://raw.githubusercontent.com/reisanar/datasets/master/fuel.csv) 

#### Loading the relevant packages and libraries:

```{r}

r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)

library(tidyverse)

install.packages("RColorBrewer")                   
library("RColorBrewer")

# Load original data
fuel_data <- read_csv("https://raw.githubusercontent.com/Dhrish96/dataviz_final_project/main/data/fuel.csv")

```


**Displaying the columns**

```{r}
str(fuel_data)
```
### The Four Main Graphs which have been visualized are as follows: 

1) Graph for the Most Popular Class of Car
2) Graph for the Most Popular Make of Car
3) Graph for Most Fuel type used in the most popular make of car
4) Graph for Annual Cost of Fuel used in the most popular make of car

**We select the required columns and create a new sub-dataset from the original dataset**

```{r}
fuel_data_new <- fuel_data %>%
  select(year, class, make, transmission, fuel_type, city_mpg_ft1, highway_mpg_ft1, combined_mpg_ft1, annual_fuel_cost_ft1)
```

*Displaying the Column Names of the new sub-dataset*
```{r}
str(fuel_data_new)
```
*Checking for any Null Values present in the sub-dataset*
```{r}
sum(is.na(fuel_data_new$year))
sum(is.na(fuel_data_new$class))
sum(is.na(fuel_data_new$make))
sum(is.na(fuel_data_new$transmission))
sum(is.na(fuel_data_new$fuel_type))
sum(is.na(fuel_data_new$city_mpg_ft1))
sum(is.na(fuel_data_new$highway_mpg_ft1))
sum(is.na(fuel_data_new$combined_mpg_ft1))
sum(is.na(fuel_data_new$annual_fuel_cost_ft1))
```
*Summarizing the columns*
```{r}
summary(fuel_data_new)
nrow(fuel_data_new)
```
```{r}
fuel_data_new <- na.omit(fuel_data_new)
sum(is.na(fuel_data_new$transmission))
```
### First Data Visualization

**Finding the Most Popular Class of Car**

```{r}
ggplot(fuel_data_new, aes(y = class)) +
  geom_bar(color="green") +
  theme_classic() +
  labs(title = "The Most Popular Class of Car",
       x = "Count",
       y = "Class of the Car",
       caption = "Source: fuel_data_new Dataset")
  

```

### Second Data Visualization
**Choosing the popular make of car and filtering it according to class of cars to plot the Histogram**


```{r}
pop_class <- c("Subcompact Cars", "Midsize Cars", "Compact Cars")

pop_class <- fuel_data_new %>% 
      filter(class %in% pop_class)

pop_class %>%
  ggplot(aes(x = make, fill=class)) +
  geom_bar(position = "stack") + 
  stat_count(geom = "text", aes(label = ..count..), vjust = 1.5, size=3) +
  theme(axis.text.x = element_text(angle = 90),legend.text = element_text(size=7), legend.position="none") +
  scale_fill_brewer(palette="Paired") +
  labs(title = "The Make of the Cars",
       subtitle ="Top three Class of Cars",
       x = "Make of the Car",
       y = "Total Count")   
```

**Counting the make of each class of car**

```{r}
pop_class %>%
  count(make, class) 
```

### Third Data Visualization
**Choosing the most used type of fuel in the car make "BMW" and plotting the Barchart**


```{r}
pop_make <- c("BMW")

pop_class_make <- pop_class %>% 
      filter(make %in% pop_make)

summary(pop_class_make)
str(pop_class_make)
unique(pop_class_make$class)

```



```{r}
ggplot(pop_class_make, aes(x = fuel_type, fill=class))+
  geom_bar()+
  coord_flip()+
  scale_fill_brewer(palette="Dark2") +
  theme_minimal()+
  theme(legend.title = element_blank(), legend.position="bottom")+
  labs(title = "Most Used Fuel Type",
       subtitle ="Top three Class of Cars of the BMW Make",
       x = "Type of Fuel",
       y = "Total Count of Cars") + 
  facet_wrap(~class)
```

### Fourth Data Visualization
**Choosing the Premium type of fuel and plotting the Boxplot for displaying the Annual cost of Fuel**


```{r}

pop_fuel <- c("Premium")

pop_cla_mak_fuel <- pop_class_make %>% 
      filter(fuel_type %in% pop_fuel)
summary(pop_cla_mak_fuel)
unique(pop_cla_mak_fuel$fuel_type)

```



```{r}
ggplot(data = pop_cla_mak_fuel) + 
  geom_boxplot(mapping=aes(y=annual_fuel_cost_ft1, group=class, color=class))+
  labs(title = "Annual Premium Fuel Cost for Cars",
       subtitle ="Top three Class of Cars of BMW Make",
       y = "Cost of Fuel in $")  

```


