---
title: "Mini Project 2"
author: "Dhrishya Menon"
date: '2022-06-14'
output: html_document
---

```{r}
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)
```

**Installing and loading packages**

```{r}

install.packages("plotly")
install.packages("RColorBrewer")
library(tidyverse)
library(ggplot2)
```
## Interactive Plot


We load the dataset
```{r}
# Load original data
birth_data <- read_csv("https://raw.githubusercontent.com/Dhrish96/dataviz_final_project/main/data/us_births_00_14.csv")
```

Displaying the column names
```{r}
colnames(birth_data)
```

```{r}
str(birth_data)
```
Displaying the first 10 data of the dataset
```{r}
head(birth_data,10)
```

*We use the group_by() function and create a new sub dataset from the original dataset and it is also summarised* 

```{r}
Year_month_grp <- birth_data %>%
  group_by(year, month) %>%
  summarise(sum_birth=sum(births))

head(Year_month_grp,10)
```


```{r}
library(RColorBrewer)
library(dplyr)
library(plotly)
```


```{r}
fig <- Year_month_grp %>%
  plot_ly(
    x = ~month, 
    y = ~sum_birth, 
    
   
    frame = ~year, 
    text = ~sum_birth, 

    hoverinfo = "text",
    hovertemplate = 'No.of Births: %{y} <extra></extra>',
    type = 'scatter',
    mode = 'lines+markers',
    marker = list(color='purple')
    
  )

fig <- fig %>% layout(
    xaxis = list(title = '<b> Months </b>'),
    yaxis = list(title = '<b> Sum of Births </b>'),
    title = "<b> Total number of Births throughout the years </b>"
    )


fig <- fig %>% animation_opts(
    1000, easing = "elastic", redraw = FALSE
  )

fig <- fig %>% animation_button(
    x = 1, xanchor = "right", y = 0, yanchor = "bottom"
  )


fig <- fig %>% animation_slider(
    currentvalue = list(prefix = "YEAR ", font = list(family = "Tahoma", color="blue"))
  )

fig
```

**We save a standalone HTML file**


```{r}
install.packages("htmlwidgets")
library(htmlwidgets)

htmlwidgets::saveWidget(
                widget = fig, #the plotly object
                file = "birth_data_figure.html", #the path & file name
                selfcontained = TRUE #creates a single html file
                )

```



## Spatial Visualization

The shapefile of the data being used is loaded

```{r}
library(sf)

loc_lakes <- "https://github.com/Dhrish96/dataviz_final_project/blob/main/data/Florida_Lakes.shp"

fla_lakes <- st_read(loc_lakes, quiet = TRUE)
```


```{r}
colnames(fla_lakes)
```

```{r}
str(fla_lakes)
```

The first 10 data in the dataset is displayed in descending order
```{r}
fla_lakes[order(-fla_lakes$SHAPEAREA), ]
```


```{r}
fla_lakes %>%
filter(COUNTY == "PALM BEACH") %>%

ggplot() +
geom_sf(aes(fill = SHAPEAREA),
alpha=0.8, col="white") +
scale_fill_viridis_c(option = 'viridis', name = "Area") +
coord_sf(crs = "+proj=merc") +  
#coord_sf(crs = "+proj=moll +ellps=WGS84") +
guides(fill=guide_legend(nrow=1)) +
theme(legend.position="bottom")+
labs(title = "Areawise distribution of Lake Okeechobee in Palm Beach County", caption="Dataset: fla_lakes")
```



## Linear Regression Model 


The dataset is grouped according to year and the number of births is summed up and stored in a new sub dataset which is used to plot the linear regression graph


```{r}
Year_grp <- birth_data %>%
  group_by(year) %>%
  summarise(sum_birth=sum(births))

head(Year_grp,10)
```


```{r}
install.packages("broom")
library(broom)
```


```{r}
linear_reg <- lm(sum_birth~year, data=Year_grp)

summary(linear_reg)

tidy(linear_reg)

glance(linear_reg)
```



```{r}
ggplot(Year_grp, aes(x = year, y = sum_birth)) +
 geom_point(color="#D55E00", size = 2) +
 geom_smooth(method = "lm", 
 formula = "y ~ x",
 color = "#009E73",
 size = 1.2,
 fill="yellow") + 
 theme_bw() +
 labs (x="Year", y="Sum of Births",
       title = "Linear Regression Model",
       caption ="Dataset : us_births_00_14") 

```

